Liveness-Based Garbage Collection for Lazy Languages

### 2015-11-29 paper/paper.tex
minefield bang, and \sc command fix
Added graphs for all BMs

\begin{center}\footnotesize
\renewcommand{\arraystretch}{1.5}
\begin{tabular}{|c|c|c|}
\hline
Premise & Transition & Rule name \\
\hline
\hline 
          & $\rho,\, (\rho', \ell, e)\!:\!S,\, \heap,\, \kappa
  \rightsquigarrow \rho',\, S,\, \heap[\ell := \kappa],\, e$    &  \sc{const}
\\
\hline
          & {$\rho, \,(\rho', \ell, e)\!:\!S,\, \heap,\, (\CONS~x~y)
\rightsquigarrow
$  $\rho',\, S,\, \heap[\ell := (\rho(x), \rho(y))],\, e$}     &  \sc{cons} \\
\hline

$\heap(\rho(x)) \mbox{ is } (v, d)$ & $\rho,\, (\rho', \ell, e)\!:\!S,\, \heap,\,
(\CAR~x)  \rightsquigarrow \rho',\, S,\, \heap[\ell := v],\, e$      &
\sc{car-select} \\
\hline

$\heap(\rho(x)) \mbox{ is } (\langle s, \rho'\rangle, d)$ & $\rho,\, S,\,  \heap,\,
(\CAR~x)  \rightsquigarrow \rho', \,(\rho, addr(\langle s, \rho'\rangle), (\CAR~x))\!:\!S,\, \heap,\, s$      &
\sc{car-1-clo} \\
\hline

$\heap(\rho(x)) \mbox{ is } \langle s, \rho'\rangle$ & $\rho,\, S,\, \heap,\, (\CAR~x)
\rightsquigarrow
\rho',\, (\rho, \rho(x), (\CAR~x))\!:\!S, \,\heap,\, s$      &
\sc{car-clo}
\\
\hline

$\heap(\rho(x)), \heap(\rho(y)) \in \mathbb{N}$
 & {$\rho,\, (\rho', \ell, e)\!:\!S,\, \heap,\, (+~x~y)  \rightsquigarrow$
$\rho', \,S,\, \heap[\ell := \heap(\rho'(x)) + \heap(\rho'(y))],\, e$}      &
\sc{prim-add} \\
\hline

$\heap(\rho(x)) \mbox{ is } \langle s, \rho'\rangle$ & $\rho,\, S,\, \heap,\, (+~x~y)
\rightsquigarrow
\rho',\, (\rho, \rho(x), (+~x~y))\!:\!S,\, \heap,\, s$      &
\sc{prim-1-clo} \\
\hline
$\heap(\rho(y)) \mbox{ is } \langle s, \rho'\rangle $ & $\rho,\, S,\, \heap,\, (+~x~y)
\rightsquigarrow
\rho', (\rho,\, \rho(y),\, (+~x~y))\!:\!S,\, \heap,\, s$      &
\sc{prim-2-clo} \\
\hline
{$\mathit{f}~\mbox{defined as}$
$~(\DEFINE~(f~\myvec{y})~e_{\mathit{f}})$}  & $\rho,\, S,\, \heap,\,
(f~\myvec{x})  \rightsquigarrow
[\myvec{y} \mapsto \rho(\myvec{x})],\, S,\, \heap,\, e_{\mathit{f}}$      &
\sc{funcall} \\
\hline
$\ell$ is a new location& {$\rho,\, S,\, \heap,\, (\LET~x\leftarrow s~\IN~e)
  \rightsquigarrow$
$\rho\oplus[x \mapsto \ell],\, S,\, \heap[\ell \mapsto \langle s,
    \lfloor\rho\rfloor_{FV(s)} %% \oplus [x \mapsto  \ell] -- NO LAZY LETS
    \rangle],\, e$} &
\sc{let} \\
\hline
$\heap(\rho(x)) \ne 0$ & $\rho,\, S,\, \heap,\, (\SIF~x~e_1~e_2)   \rightsquigarrow
\rho,\, S,\, \heap,\,  e_1$ & \sc{if-true} \\
\hline
$\heap(\rho(x)) = 0$ & $\rho,\, S,\, \heap,\, (\SIF~x~e_1~e_2)   \rightsquigarrow
\rho,\, S, \,\heap, \, e_2$ & \sc{if-false} \\
\hline
$\heap(\rho(x)) = \langle s, \rho' \rangle $ & {$\rho,\, S,\, \heap,\,
  (\SIF~x~e_1~e_2)   \rightsquigarrow
\rho',\, (\rho, \rho(x), (\SIF~x~e_1~e_2))\!:\!S,\, \heap, \, s$}
&
\sc{if-clo} \\
\hline
{$\heap(\rho(x))~\mbox{is in WHNF with value}~v$}& $\rho,\, (\rho', \ell,
e)\!:\!S,\, \heap,\,
(\SRETURN~x)  \rightsquigarrow \rho',\, S,\, \heap[\ell := v],\, e$ &
\sc{return-whnf}\\
\hline
$\heap(\rho(x)) = \langle s, \rho' \rangle $ & {$\rho,\, S,\, \heap,\, (\SRETURN~x)
  \rightsquigarrow$
$\rho',\, (\rho, \rho(x), (\SRETURN~x))\!:\!S,\, \heap,\,  s$} &
\sc{return-clo} \\
\hline
\end{tabular}
\caption{The small-step semantics for the language. \label{fig:lang-semantics}}
\end{center}
%}\comment
\end{figure*}

@@

%% \hline
%% \makecell[t]{$GC(\rho_1, S, \heap_1, \sigma) = (\rho, S, \heap)$,}\\ $\heap(\rho(x)) \mbox{ is } \bot$ & $\rho, S,
%%   \heap, (\CAR~x), \cred{\sigma} \rightsquigarrow \bang$   &
%% \sc{car-clo-bang}
%% \\
%% \hline
%% \makecell[t]{$GC(\rho_1, S, \heap_1, \Lv, \sigma) = (\rho, S, \heap)$,}\\ $\heap(\rho(x))
%% \mbox{ is } \langle s, \rho'\rangle$ & $\rho, S,
%%   \heap, (\CAR~x), \cred{\sigma} \rightsquigarrow   \rho', (\rho, x,
%%   (\CAR~x), \cred{\sigma})\!:\!S, \heap, s, \cred{(\clazy \cup \acar)\sigma }$         &
%% \sc{car-clo}\\
%% \hline
%% \end{tabular}

@@

%%           &\makecell{ $\rho, (\rho', x, e, \cred{\sigma'})\!:\!S,
%%   \heap, \kappa, \cred{\sigma}$  $\rightsquigarrow \rho', S, \heap[\rho'(x) :=
%%     \kappa], e, \cred{\sigma'}$ }   &  \sc{const}
%% \\
%% \hline
%%           & \makecell{$\rho, (\rho', z, e, \cred{\sigma'})\!:\!S, \heap, (\CONS~x~y), \cred{\sigma}$  $\rightsquigarrow
%%   \rho', S, \heap[\rho'(z) := (\rho(x),\rho(y))], e, \cred{\sigma'}$}     &  \sc{cons} \\
%% \hline
%% $\heap(\rho(x)) \mbox{ is } (\ell_1, \ell_2)$ & \makecell{$\rho, (\rho', z, e,
%%   \cred{\sigma'} )\!:\!S, \heap, (\CAR~x), \cred{\sigma}$  $
%%   \rightsquigarrow \rho', S, \heap[\rho'(z) := \heap(\ell_1)], e, \cred{\sigma'}$}      &
%% \sc{car-whnf} \\
%% \hline
%% $\heap(\rho(x)) \mbox{ is } \langle s, \rho'\rangle$ &\makecell{ $\rho, S,
%%   \heap, (\CAR~x), \cred{\sigma}$  $\rightsquigarrow \rho', (\rho, x,
%%   (\CAR~x), \cred{\sigma})\!:\!S, \heap, s, \cred{(\clazy \cup \acar)\sigma }$}      &
%% \sc{car-clo}
%% \\

%% \hline
%% $\heap(\rho(x)) \mbox{ is } (\langle s, \rho'\rangle, d)$ & $\rho,\, S,\,  \heap,\,
%% (\CAR~x), \cred{\sigma} \rightsquigarrow \rho', \,(\rho, addr(\langle
%% s, \rho'\rangle), (\CAR~x),\cred{\sigma} )\!:\!S,\, \heap,\, s, \, \cred{\sigma}$      &
%% \sc{car-1-clo} \\

%% \hline
%% $\heap(\rho(x)), \heap(\rho(y)) \in \mathbb{N}$
%%  & \makecell{$\rho, (\rho', z, e, \cred{\sigma'})\!:\!S, \heap,
%%   (+~x~y), \cred{\sigma}$   $\rightsquigarrow \rho', S, \heap[\rho'(z)
%%     := \heap(\rho'(x)) + \heap(\rho'(y))], e, \cred{\sigma'}$}      &
%% \sc{prim-whnf} \\
%% \hline
%% $\heap(\rho(x)) \mbox{ is } \langle s, \rho'\rangle$ &\makecell{$\rho, S,
%%   \heap, (+~x~y), \cred{\sigma}$  $\rightsquigarrow \rho', (\rho, x,
%%   (+~x~y), \cred{\sigma})\!:\!S, \heap, s, \cred{\clazy\sigma}$}      &
%% \sc{prim-1-clo} \\
%% \hline
%% $\heap(\rho(y)) \mbox{ is } \langle s, \rho'\rangle $ & \makecell{$\rho,
%%   S, \heap, (+~x~y), \cred{\sigma}$  $\rightsquigarrow \rho', (\rho, y,
%%   (+~x~y), \cred{\sigma})\!:\!S, \heap, s, \cred{\clazy\sigma}$}      &
%% \sc{prim-2-clo} \\
%% \hline
%% {$\mathit{g}~\mbox{defined as}$
%% $~(\DEFINE~(g~\myvec{y})~e_{\mathit{g}})$}  & \makecell{$\rho, S, \heap,
%%   (g~\myvec{x}), \cred{\sigma}$  $\rightsquigarrow [\myvec{y} \mapsto
%%     \rho(\myvec{x})], S, \heap, e_{\mathit{g}}, \cred{\sigma}$}      &
%% \sc{funcall} \\
%% \hline
%% $\ell$ is a new location& \makecell{$\rho, S, \heap, (\LET~x\leftarrow
%%   s~\IN~e), \cred{\sigma}$  $ \rightsquigarrow \rho\oplus[x \mapsto \ell], S, \heap[\ell \mapsto \langle s, \lfloor\rho\rfloor_{FV(s)}  \oplus [x \mapsto
%%   \ell]\rangle], e, \cred{\sigma}$} &
%% \sc{let} \\ 
%% \hline
%% $\heap(\rho(x)) \ne 0$ & \makecell{$\rho, S, \heap, (\pi:\SIF~\psi:x~e_1~e_2),
%%   \cred{\sigma}$  $\rightsquigarrow \rho, S, \heap,  e_1, \cred{\sigma}$} & \sc{if-true} \\
%% \hline
%% $\heap(\rho(x)) = 0$ & \makecell{$\rho, S, \heap, (\pi:\SIF~\psi:x~e_1~e_2),
%%   \cred{\sigma}$   $\rightsquigarrow
%% \rho, S, \heap,  e_2, \cred{\sigma}$} & \sc{if-false} \\
%% \hline
%% $\heap(\rho(x)) = \langle s, \rho' \rangle $ & \makecell{$\rho, S, \heap,
%%   (\pi:\SIF~\psi:x~e_1~e_2), \cred{\sigma}$ $\rightsquigarrow
%% \rho', (\rho, x, (\SIF~x~e_1~e_2),  \cred{\sigma})\!:\!S, \heap, s,
%% \cred{\clazy\sigma}$}
%% &
%% \sc{if-clo} \\
%% \hline
%% {$\heap(\rho(x))~\mbox{is}$ $\mbox{whnf with value}~v$}& \makecell{$\rho,
%%   (\rho', z, e, \cred{\sigma'})\!:\!S, \heap, (\SRETURN~x), \cred{\sigma}$  $\rightsquigarrow \rho', S, \heap[\rho'(z) := v], e,
%%   \cred{\sigma'}$} &
%% \sc{return-whnf}\\
%% \hline
%% $\heap(\rho(x)) = \langle s, \rho' \rangle $ & \makecell{$\rho, S, \heap,
%%   (\psi:\SRETURN~x), \cred{\sigma}$  $
%%   \rightsquigarrow$
%% $\rho',~ (\rho, x, (\SRETURN~x), \cred{\sigma})\!:\!S, \heap,  s,
%%   \cred{\sigma}$} &
%% \sc{return-clo} \\
%% \hline
%% \end{tabular}
%% \caption{Minefield semantics.\label{fig:minefield-semantics}}
%% \end{center}
%% \vskip -5mm
%% \end{figure*}

@@

\hline
\makecell[t]{$\cred{GC(\rho_1, S, \heap_1, (\CAR~x), \sigma) = (\rho, S, \heap)}$,} $\heap(\rho(x)) \mbox{ is } \bot$ & $\rho, S,
  \heap, (\CAR~x), \cred{\sigma} \rightsquigarrow \bang$   &
\sc{car-clo-bang} 
\\
\hline
\makecell[t]{$\cred{GC(\rho_1, S, \heap_1, (\CAR~x), \sigma) = (\rho, S, \heap)}$,} $\heap(\rho(x))
\mbox{ is } \langle s, \rho'\rangle$ & $\rho, S,
  \heap, (\CAR~x), \cred{\sigma} \rightsquigarrow   \rho', (\rho, x,
  (\CAR~x), \cred{\sigma})\!:\!S, \heap, s, \cred{(\clazy \cup \acar)\sigma }$         &
\sc{car-clo}\\
\hline
\end{tabular}
\end{center}
\caption{Minefield semantics for \sc{car-clo-bang} and
  \sc{car-clo-bang} \label{fig:minefield-semantics-for-some}}
\end{figure*}

@@

\newcommand{\hgt}{2.9cm}
\begin{figure*}[t]
\renewcommand{\arraystretch}{.1}
\begin{tabular}{@{}c@{}@{}c@{}@{}c@{}}
   \hskip -4mm{\epsfig{file=fibheap_win.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=sudoku_win.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=nperm_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=paraffins_win.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=lcss_win.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=huffman_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=knightstour_win.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=nqueens_win.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=deriv_win.eps, height=\hgt}}
%\\ (a) & (b)
\end{tabular}%\vskip -10mm
 \caption{Memory usage.  
%% Column (a) shows complete  usage, column 
%%    (b) enlarges  a part  of the  usage.  
The blue and the red curves indicate the number of cons
cells  in  the  active   semi-space  for  RGC  and  LGC
respectively.  The black curve represents the number of
reachable cells and the  lightblue curve represents the
number  of  cells  that  are actually  live  (of  which
liveness analysis does a static approximation).  x-axis
is the time measured  in number of cons-cells allocated
(scaled down by factor $10^5$). y-axis is the number of
cons-cells (scaled down by $10^3$).}
\label{fig:memory-usage} \figrule
\end{figure*}

@@

\begin{tabular}{|c|c|c|}
\hline
Premise & Transition & Rule name \\ 
\hline
\hline
&\makecell{ $\rho, (\rho', \ell, e, \cred{\sigma'})\!:\!S,
  \heap, \kappa, \cred{\sigma}$  $\rightsquigarrow \rho', S, \heap[\ell :=
    \kappa], e, \cred{\sigma'}$ }   &  \sc{const}
\\
\hline
& \makecell{$\rho, (\rho', \ell, e, \cred{\sigma'})\!:\!S, \heap, (\CONS~x~y), \cred{\sigma}$  $\rightsquigarrow
  \rho', S, \heap[\ell := (\rho(x),\rho(y))], e, \cred{\sigma'}$}     &  \sc{cons} \\
\hline
$\heap(\rho(x)) \mbox{ is } (v, d)$ & \makecell{$\rho, (\rho', \ell, e,
  \cred{\sigma'} )\!:\!S, \heap, (\CAR~x), \cred{\sigma}$  $
  \rightsquigarrow \rho', S, \heap[\ell := v], e, \cred{\sigma'}$}      &
\sc{car-select} \\
\hline
$\heap(\rho(x)) \mbox{ is } (\langle s, \rho'\rangle, d)$ & $\rho,\, S,\,  \heap,\,
(\CAR~x), \cred{\sigma} \rightsquigarrow \rho', \,(\rho, addr(\langle
s, \rho'\rangle), (\CAR~x),\cred{\sigma} )\!:\!S,\, \heap,\, s, \, \cred{\sigma}$      &
\sc{car-1-clo} \\

@@

$\heap(\rho(x)) \mbox{ is } \langle s, \rho'\rangle$ &\makecell{ $\rho, S,
  \heap, (\CAR~x), \cred{\sigma}$  $\rightsquigarrow \rho', (\rho, x,
  (\CAR~x), \cred{\sigma})\!:\!S, \heap, s, \cred{(\clazy \cup \acar)\sigma }$}      &
\sc{car-clo}
\\

>>>>

\begin{center}\footnotesize
\renewcommand{\arraystretch}{1.5}
\begin{tabular}{|c|c|c|}
\hline
Premise & Transition & Rule name \\
\hline
\hline 
          & $\rho,\, (\rho', \ell, e)\!:\!S,\, \heap,\, \kappa
  \rightsquigarrow \rho',\, S,\, \heap[\ell := \kappa],\, e$    &  {\sc const}
\\
\hline
          & {$\rho, \,(\rho', \ell, e)\!:\!S,\, \heap,\, (\CONS~x~y)
\rightsquigarrow
$  $\rho',\, S,\, \heap[\ell := (\rho(x), \rho(y))],\, e$}     &  {\sc cons} \\
\hline

$\heap(\rho(x)) \mbox{ is } (v, d)$ & $\rho,\, (\rho', \ell, e)\!:\!S,\, \heap,\,
(\CAR~x)  \rightsquigarrow \rho',\, S,\, \heap[\ell := v],\, e$      &
{\sc car-select} \\
\hline

$\heap(\rho(x)) \mbox{ is } (\langle s, \rho'\rangle, d)$ & $\rho,\, S,\,  \heap,\,
(\CAR~x)  \rightsquigarrow \rho', \,(\rho, addr(\langle s, \rho'\rangle), (\CAR~x))\!:\!S,\, \heap,\, s$      &
{\sc car-1-clo} \\
\hline

$\heap(\rho(x)) \mbox{ is } \langle s, \rho'\rangle$ & $\rho,\, S,\, \heap,\, (\CAR~x)
\rightsquigarrow
\rho',\, (\rho, \rho(x), (\CAR~x))\!:\!S, \,\heap,\, s$      &
{\sc car-clo}
\\
\hline

$\heap(\rho(x)), \heap(\rho(y)) \in \mathbb{N}$
 & {$\rho,\, (\rho', \ell, e)\!:\!S,\, \heap,\, (+~x~y)  \rightsquigarrow$
$\rho', \,S,\, \heap[\ell := \heap(\rho'(x)) + \heap(\rho'(y))],\, e$}      &
{\sc prim-add} \\
\hline

$\heap(\rho(x)) \mbox{ is } \langle s, \rho'\rangle$ & $\rho,\, S,\, \heap,\, (+~x~y)
\rightsquigarrow
\rho',\, (\rho, \rho(x), (+~x~y))\!:\!S,\, \heap,\, s$      &
{\sc prim-1-clo} \\
\hline
$\heap(\rho(y)) \mbox{ is } \langle s, \rho'\rangle $ & $\rho,\, S,\, \heap,\, (+~x~y)
\rightsquigarrow
\rho', (\rho,\, \rho(y),\, (+~x~y))\!:\!S,\, \heap,\, s$      &
{\sc prim-2-clo} \\
\hline
{$\mathit{f}~\mbox{defined as}$
$~(\DEFINE~(f~\myvec{y})~e_{\mathit{f}})$}  & $\rho,\, S,\, \heap,\,
(f~\myvec{x})  \rightsquigarrow
[\myvec{y} \mapsto \rho(\myvec{x})],\, S,\, \heap,\, e_{\mathit{f}}$      &
{\sc funcall} \\
\hline
$\ell$ is a new location& {$\rho,\, S,\, \heap,\, (\LET~x\leftarrow s~\IN~e)
  \rightsquigarrow$
$\rho\oplus[x \mapsto \ell],\, S,\, \heap[\ell \mapsto \langle s,
    \lfloor\rho\rfloor_{FV(s)} %% \oplus [x \mapsto  \ell] -- NO LAZY LETS
    \rangle],\, e$} &
{\sc let} \\
\hline
$\heap(\rho(x)) \ne 0$ & $\rho,\, S,\, \heap,\, (\SIF~x~e_1~e_2)   \rightsquigarrow
\rho,\, S,\, \heap,\,  e_1$ & {\sc if-true} \\
\hline
$\heap(\rho(x)) = 0$ & $\rho,\, S,\, \heap,\, (\SIF~x~e_1~e_2)   \rightsquigarrow
\rho,\, S, \,\heap, \, e_2$ & {\sc if-false} \\
\hline
$\heap(\rho(x)) = \langle s, \rho' \rangle $ & {$\rho,\, S,\, \heap,\,
  (\SIF~x~e_1~e_2)   \rightsquigarrow
\rho',\, (\rho, \rho(x), (\SIF~x~e_1~e_2))\!:\!S,\, \heap, \, s$}
&
{\sc if-clo} \\
\hline
{$\heap(\rho(x))~\mbox{is in WHNF with value}~v$}& $\rho,\, (\rho', \ell,
e)\!:\!S,\, \heap,\,
(\SRETURN~x)  \rightsquigarrow \rho',\, S,\, \heap[\ell := v],\, e$ &
{\sc return-whnf}\\
\hline
$\heap(\rho(x)) = \langle s, \rho' \rangle $ & {$\rho,\, S,\, \heap,\, (\SRETURN~x)
  \rightsquigarrow$
$\rho',\, (\rho, \rho(x), (\SRETURN~x))\!:\!S,\, \heap,\,  s$} &
{\sc return-clo} \\
\hline
\end{tabular}
\caption{The small-step semantics for the language. \label{fig:lang-semantics}}
\end{center}
%}\comment
\end{figure*}

@@

%% \hline
%% \makecell[t]{$GC(\rho_1, S, \heap_1, \sigma) = (\rho, S, \heap)$,}\\ $\heap(\rho(x)) \mbox{ is } \bot$ & $\rho, S,
%%   \heap, (\CAR~x), \cred{\sigma} \rightsquigarrow \bang$   &
%% {\sc car-clo-bang}
%% \\
%% \hline
%% \makecell[t]{$GC(\rho_1, S, \heap_1, \Lv, \sigma) = (\rho, S, \heap)$,}\\ $\heap(\rho(x))
%% \mbox{ is } \langle s, \rho'\rangle$ & $\rho, S,
%%   \heap, (\CAR~x), \cred{\sigma} \rightsquigarrow   \rho', (\rho, x,
%%   (\CAR~x), \cred{\sigma})\!:\!S, \heap, s, \cred{(\clazy \cup \acar)\sigma }$         &
%% {\sc car-clo}\\
%% \hline
%% \end{tabular}

@@

%%           &\makecell{ $\rho, (\rho', x, e, \cred{\sigma'})\!:\!S,
%%   \heap, \kappa, \cred{\sigma}$  $\rightsquigarrow \rho', S, \heap[\rho'(x) :=
%%     \kappa], e, \cred{\sigma'}$ }   &  {\sc const}
%% \\
%% \hline
%%           & \makecell{$\rho, (\rho', z, e, \cred{\sigma'})\!:\!S, \heap, (\CONS~x~y), \cred{\sigma}$  $\rightsquigarrow
%%   \rho', S, \heap[\rho'(z) := (\rho(x),\rho(y))], e, \cred{\sigma'}$}     &  {\sc cons} \\
%% \hline
%% $\heap(\rho(x)) \mbox{ is } (\ell_1, \ell_2)$ & \makecell{$\rho, (\rho', z, e,
%%   \cred{\sigma'} )\!:\!S, \heap, (\CAR~x), \cred{\sigma}$  $
%%   \rightsquigarrow \rho', S, \heap[\rho'(z) := \heap(\ell_1)], e, \cred{\sigma'}$}      &
%% {\sc car-whnf} \\
%% \hline
%% $\heap(\rho(x)) \mbox{ is } \langle s, \rho'\rangle$ &\makecell{ $\rho, S,
%%   \heap, (\CAR~x), \cred{\sigma}$  $\rightsquigarrow \rho', (\rho, x,
%%   (\CAR~x), \cred{\sigma})\!:\!S, \heap, s, \cred{(\clazy \cup \acar)\sigma }$}      &
%% {\sc car-clo}
%% \\

%% \hline
%% $\heap(\rho(x)) \mbox{ is } (\langle s, \rho'\rangle, d)$ & $\rho,\, S,\,  \heap,\,
%% (\CAR~x), \cred{\sigma} \rightsquigarrow \rho', \,(\rho, addr(\langle
%% s, \rho'\rangle), (\CAR~x),\cred{\sigma} )\!:\!S,\, \heap,\, s, \, \cred{\sigma}$      &
%% {\sc car-1-clo} \\

%% \hline
%% $\heap(\rho(x)), \heap(\rho(y)) \in \mathbb{N}$
%%  & \makecell{$\rho, (\rho', z, e, \cred{\sigma'})\!:\!S, \heap,
%%   (+~x~y), \cred{\sigma}$   $\rightsquigarrow \rho', S, \heap[\rho'(z)
%%     := \heap(\rho'(x)) + \heap(\rho'(y))], e, \cred{\sigma'}$}      &
%% {\sc prim-whnf} \\
%% \hline
%% $\heap(\rho(x)) \mbox{ is } \langle s, \rho'\rangle$ &\makecell{$\rho, S,
%%   \heap, (+~x~y), \cred{\sigma}$  $\rightsquigarrow \rho', (\rho, x,
%%   (+~x~y), \cred{\sigma})\!:\!S, \heap, s, \cred{\clazy\sigma}$}      &
%% {\sc prim-1-clo} \\
%% \hline
%% $\heap(\rho(y)) \mbox{ is } \langle s, \rho'\rangle $ & \makecell{$\rho,
%%   S, \heap, (+~x~y), \cred{\sigma}$  $\rightsquigarrow \rho', (\rho, y,
%%   (+~x~y), \cred{\sigma})\!:\!S, \heap, s, \cred{\clazy\sigma}$}      &
%% {\sc prim-2-clo} \\
%% \hline
%% {$\mathit{g}~\mbox{defined as}$
%% $~(\DEFINE~(g~\myvec{y})~e_{\mathit{g}})$}  & \makecell{$\rho, S, \heap,
%%   (g~\myvec{x}), \cred{\sigma}$  $\rightsquigarrow [\myvec{y} \mapsto
%%     \rho(\myvec{x})], S, \heap, e_{\mathit{g}}, \cred{\sigma}$}      &
%% {\sc funcall} \\
%% \hline
%% $\ell$ is a new location& \makecell{$\rho, S, \heap, (\LET~x\leftarrow
%%   s~\IN~e), \cred{\sigma}$  $ \rightsquigarrow \rho\oplus[x \mapsto \ell], S, \heap[\ell \mapsto \langle s, \lfloor\rho\rfloor_{FV(s)}  \oplus [x \mapsto
%%   \ell]\rangle], e, \cred{\sigma}$} &
%% {\sc let} \\ 
%% \hline
%% $\heap(\rho(x)) \ne 0$ & \makecell{$\rho, S, \heap, (\pi:\SIF~\psi:x~e_1~e_2),
%%   \cred{\sigma}$  $\rightsquigarrow \rho, S, \heap,  e_1, \cred{\sigma}$} & {\sc if-true} \\
%% \hline
%% $\heap(\rho(x)) = 0$ & \makecell{$\rho, S, \heap, (\pi:\SIF~\psi:x~e_1~e_2),
%%   \cred{\sigma}$   $\rightsquigarrow
%% \rho, S, \heap,  e_2, \cred{\sigma}$} & {\sc if-false} \\
%% \hline
%% $\heap(\rho(x)) = \langle s, \rho' \rangle $ & \makecell{$\rho, S, \heap,
%%   (\pi:\SIF~\psi:x~e_1~e_2), \cred{\sigma}$ $\rightsquigarrow
%% \rho', (\rho, x, (\SIF~x~e_1~e_2),  \cred{\sigma})\!:\!S, \heap, s,
%% \cred{\clazy\sigma}$}
%% &
%% {\sc if-clo} \\
%% \hline
%% {$\heap(\rho(x))~\mbox{is}$ $\mbox{whnf with value}~v$}& \makecell{$\rho,
%%   (\rho', z, e, \cred{\sigma'})\!:\!S, \heap, (\SRETURN~x), \cred{\sigma}$  $\rightsquigarrow \rho', S, \heap[\rho'(z) := v], e,
%%   \cred{\sigma'}$} &
%% {\sc return-whnf}\\
%% \hline
%% $\heap(\rho(x)) = \langle s, \rho' \rangle $ & \makecell{$\rho, S, \heap,
%%   (\psi:\SRETURN~x), \cred{\sigma}$  $
%%   \rightsquigarrow$
%% $\rho',~ (\rho, x, (\SRETURN~x), \cred{\sigma})\!:\!S, \heap,  s,
%%   \cred{\sigma}$} &
%% {\sc return-clo} \\
%% \hline
%% \end{tabular}
%% \caption{Minefield semantics.\label{fig:minefield-semantics}}
%% \end{center}
%% \vskip -5mm
%% \end{figure*}

@@

\hline
\makecell[t]{$\cred{GC(\rho_1, S, \heap_1, (\CAR~x), \sigma) = (\rho, S, \heap)}$,} $\heap(\rho(x)) \mbox{ is } \bot$ & $\rho, S,
  \heap, (\CAR~x), \cred{\sigma} \rightsquigarrow \bang$   &
{\sc car-clo-bang} 
\\
\hline
\makecell[t]{$\cred{GC(\rho_1, S, \heap_1, (\CAR~x), \sigma) = (\rho, S, \heap)}$,} $\heap(\rho(x))
\mbox{ is } \langle s, \rho'\rangle$ & $\rho, S,
  \heap, (\CAR~x), \cred{\sigma} \rightsquigarrow   \rho', (\rho, \ell,
  (\CAR~x), \cred{\sigma})\!:\!S, \heap, s, \cred{(\clazy \cup \acar)\sigma }$         &
{\sc car-clo}\\
\hline
\end{tabular}
\end{center}
\caption{Minefield semantics for {\sc car-clo-bang} and
  {\sc car-clo} \label{fig:minefield-semantics-for-some}}
\end{figure*}

@@

\newcommand{\hgt}{2.9cm}
\begin{figure*}[t]
\renewcommand{\arraystretch}{.1}
\begin{tabular}{@{}c@{}@{}c@{}@{}c@{}}
   \hskip -4mm{\epsfig{file=fibheap_win.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=sudoku_win.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=nperm_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=paraffins_win.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=lcss_win.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=huffman_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=knightstour_win.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=nqueens_win.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=deriv_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=treejoin.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=gc_bench_win.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=lambda_win.eps, height=\hgt}}
%\\ (a) & (b)
\end{tabular}%\vskip -10mm
 \caption{Memory usage.  
%% Column (a) shows complete  usage, column 
%%    (b) enlarges  a part  of the  usage.  
The blue and the red curves indicate the number of cons
cells  in  the  active   semi-space  for  RGC  and  LGC
respectively.  The black curve represents the number of
reachable cells and the  lightblue curve represents the
number  of  cells  that  are actually  live  (of  which
liveness analysis does a static approximation).  x-axis
is the time measured  in number of cons-cells allocated
(scaled down by factor $10^5$). y-axis is the number of
cons-cells (scaled down by $10^3$).}
\label{fig:memory-usage} \figrule
\end{figure*}

@@

\begin{tabular}{|c|c|c|}
\hline
Premise & Transition & Rule name \\ 
\hline
\hline
&\makecell{ $\rho, (\rho', \ell, e, \cred{\sigma'})\!:\!S,
  \heap, \kappa, \cred{\sigma}$  $\rightsquigarrow \rho', S, \heap[\ell :=
    \kappa], e, \cred{\sigma'}$ }   &  {\sc const}
\\
\hline
& \makecell{$\rho, (\rho', \ell, e, \cred{\sigma'})\!:\!S, \heap, (\CONS~x~y), \cred{\sigma}$  $\rightsquigarrow
  \rho', S, \heap[\ell := (\rho(x),\rho(y))], e, \cred{\sigma'}$}     &  {\sc cons} \\
\hline
$\heap(\rho(x)) \mbox{ is } (v, d)$ & \makecell{$\rho, (\rho', \ell, e,
  \cred{\sigma'} )\!:\!S, \heap, (\CAR~x), \cred{\sigma}$  $
  \rightsquigarrow \rho', S, \heap[\ell := v], e, \cred{\sigma'}$}      &
{\sc car-select} \\
\hline
$\heap(\rho(x)) \mbox{ is } (\langle s, \rho'\rangle, d)$ & $\rho,\, S,\,  \heap,\,
(\CAR~x), \cred{\sigma} \rightsquigarrow \rho', \,(\rho, addr(\langle
s, \rho'\rangle), (\CAR~x),\cred{\sigma} )\!:\!S,\, \heap,\, s, \, \cred{\sigma}$      &
{\sc car-1-clo} \\

@@

$\heap(\rho(x)) \mbox{ is } \langle s, \rho'\rangle$ &\makecell{ $\rho, S,
  \heap, (\CAR~x), \cred{\sigma}$  $\rightsquigarrow \rho', (\rho, x,
  (\CAR~x), \cred{\sigma})\!:\!S, \heap, s, \cred{(\clazy \cup \acar)\sigma }$}      &
{\sc car-clo}
\\
